### 2014-04-16 cs250-hw/hw2.tex
c2

\documentclass{article}
\usepackage{geometry}
\usepackage[namelimits,sumlimits]{amsmath}
\usepackage{amssymb,amsfonts}
\usepackage{multicol}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage[cm]{fullpage}
\newcommand{\tab}{\hspace*{5em}}
\newcommand{\conj}{\overline}
\newcommand{\dd}{\partial}
\newcommand{\ep}{\epsilon}
\newcommand{\openm}{\begin{pmatrix}}
\newcommand{\closem}{\end{pmatrix}}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\im}{im}
\DeclareMathOperator{\Span}{span}
\DeclareMathOperator{\Null}{null}
\newcommand{\nc}{\newcommand}
\newcommand{\rn}{\mathbb{R}}
\newcommand{\zn}{\mathbb{Z}}
\nc{\cn}{\mathbb{C}}
\nc{\ssn}[1]{\subsubsection*{#1}}
\nc{\inner}[2]{\langle #1,#2\rangle}
\nc{\h}[1]{\widehat{#1}}
\nc{\tl}[1]{\widetilde{#1}}
\nc{\norm}[1]{\left\|{#1}\right\|}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
\begin{document}
Name: Hall Liu

Date: \today 
\vspace{1.5cm}

\subsection*{1}
\ssn{a}
First, note that since $F$ is strictly increasing, we must have that $F(x)\in(0,1)$ for any $x\in\rn$ and $F^{-1}$ exists. Then, for any $y\in(0,1)$, we have $P(Y<y)=P(F(X)<y)=P(X<F^{-1}(y))=F(F^{-1}(y))=y$, which implies that $Y$ has a cdf of $0$ on $(-\infty, 0)$, $y$ on $[0,1]$, and $1$ on $(1, \infty)$. Thus, we have $Y\sim\text{Unif}(0, 1)$.

Going the other way, we have $P(X<x)=P(F^{-1}(U)<x)=P(U<F(x))=F(x)$.

\ssn{b}
The joint density $f_{X,Y}(x,y)$ of $X$ and $Y$ is the product of their densities (which is 1) since they are independent. Thus, for the difference, for $z\in(0,1]$, we have
\[P(Z<z)=P(X-Y<z)=P(X<z+Y)=\int_0^{1-z}\int_0^{z+y}1dxdy+\int_{1-z}^1\int_0^11dxdy=\int_0^{1-z}(z+y)dy+z=-\frac{z^2}{2}+z+\frac{1}{2}\]
For $z\in[-1,0]$, we have
\[P(Z<z)=P(X-Y<z)=P(X<z+Y)=\int_{-z}^1\int_0^{y+z}1dxdy=\int_{-z}^1(z+y)dy=\frac{z^2}{2}+z+\frac{1}{2}\]
Differentiating produces the density $z+1$ on $[-1,0]$ and $-z+1$ on $(0,1]$.

For the minimum, we want $X<z$ or $Y<z$, which is equivalent to the negation of ($X\geq z$ and $Y\geq z$), so the integral for the cdf of $Z$ is 
\[1-\int_{1-z}^1\int_{1-z}^11dxdy=1-(1-z)^2\]
and differentiating this produces the density $2(1-z)$.
\ssn{c}
The density of $X$ is $f_X(x)=\frac{1}{\sqrt{2\pi}}e^{-\frac{x^2}{2}}$. Taking the expectation of $e^X$ wrt this density gives
\[\frac{1}{\sqrt{2\pi}}\int_{-\infty}^\infty\exp\left(-\frac{x^2}{2}+x\right)dx=\frac{1}{\sqrt{2\pi}}e^{1/2}\int_{-\infty}^\infty\exp\left(-\frac{(x-1)^2}{2}\right)=e^{1/2}\]
The variance can be calculated by $E(Y^2)-(E(Y))^2$, and evaluating $E(Y^2)=E(e^{2X})$ gives
\[\frac{1}{\sqrt{2\pi}}\int_{-\infty}^\infty\exp\left(-\frac{x^2}{2}+2x\right)dx=\frac{1}{\sqrt{2\pi}}e^2\int_{-\infty}^\infty\exp\left(-\frac{(x-2)^2}{2}\right)=e^2\]
so the variance is $e^2-e$
\ssn{d}
$\var_Y(Y|X)=E_Y(Y^2|X)-(E_Y(Y|X))^2$, so the first term is $E_X(\var_Y(Y|X))=E_X(E_Y(Y^2|X))-E_X((E_Y(Y|X))^2)=E_Y(Y^2)-E_X((E_Y(Y|X))^2)$. Similarly, the second term is $E_X((E_Y(Y|X))^2)-(E_X(E_Y(Y|X)))^2=E_X((E_Y(Y|X))^2)-(E_Y(Y))^2$. Adding them together cancels out the $E_X((E_Y(Y|X))^2)$ term, which results in $E_Y(Y^2)-(E_Y(Y))^2=\var(Y)$.
\subsection*{2}
\ssn{a}
Since $X^TX$ is nonsingular and $n>d$, this means that $X$ has rank $d$. The least-squares estimate $\h{\beta}$ satisfies $\h{\beta}=\text{argmin}_\beta\|X\beta-y\|_2^2$. Decompose $X$ via the SVD $X=U\Sigma V^T$. Then, we have 
\[\|X\beta-y\|_2^2=\norm{U\Sigma V^T\beta-y}_2^2=\norm{\Sigma \gamma-U^Ty}_2^2\]
where $\gamma=V^T\beta$, by the unitary invariance of the $2$-norm. Let $z=U^Ty$. We know that precisely the first $d$ entries in $\Sigma $ are nonzero because $X$ is rank $d$, so the norm expression is 
\[\sum_{i=1}^d(\sigma_i\gamma_i-z_i)^2+\sum_{i=d+1}^nz_i^2\]
To minimize this, we set $\gamma_i=z_i/\sigma_i$ for $i\in[1..d]$, and we can choose the others arbitrarily. Might as well make the others zero, so then we have that $\h{\beta}=V\h{\gamma}=V\Sigma^+U^Ty$, where $\Sigma^+$ is diagonal with entries $\frac{1}{\sigma_i}$ where $\sigma_i\neq0$ and zero elsewhere.

Now, note that both $V\Sigma^+U^TX$ and $(X^TX)^{-1}X^T$ satisfy the requirements for being the Moore-Penrose pseudoinverse of $X$. The calculations are a bit tedious to type out and not very illuminating, so I'll omit them. Thus, by the uniqueness of the pseudoinverse, they must be equal. Thus, we have $\h{\beta}=(X^TX)^{-1}X^Ty$.
\ssn{b}
$HX=X(X^TX)^{-1}X^TX=X$
\ssn{c}
$H^T=(X(X^TX)^{-1}X^T)^T=X(X^TX)^{-T}X^T=X(X^TX)^{-1}X^T$
\ssn{d}
$HH=X(X^TX)^{-1}X^TX(X^TX)^{-1}X^T=X(X^TX)^{-1}X^T=H$
\ssn{e}
By (d), $H$ is a projection matrix. Thus, we just need to show that the image of $H$ is the column space. Since the image of $X$ is the column space (a subset of $\rn^n$), this means that we need to show that $(X^TX)^{-1}X^T$ has rank $d$ (or that its image is all of $\rn^d$). This is true because $(X^TX)^{-1}$ is invertible and $X^T$ has rank $d$ also. 
\ssn{f}
Traces of products are invariant under cyclic permutation, so $\tr(H)=\tr(X(X^TX)^{-1}X^T)=\tr(X^TX(X^TX)^{-1})=\tr(I_d)=d$.
\subsection*{3}
\ssn{b}
I have a general policy against running unknown scripts on my computer, and since I know that I have a large part of the Python toolchain set up already, I decided to read the setup script (\verb|toolkit.sh|) and replicate the necessary parts.

The first difficulty came when trying to install the packages. I ran into a syntax error while installing czipfile via pip, and this turned out to be because my system defaults to Python 3. Deleting the virtualenv and re-running mkvirtualenv with the option \verb|--python=/usr/bin/python2.7| fixed the issue.

The Python version problem came up again when creating the bin/notebook file. I modified the line to start with \verb|python2.7 /usr/bin/ipython| to force the script to use 
\end{document}

>>>>

\documentclass{article}
\usepackage{geometry}
\usepackage[namelimits,sumlimits]{amsmath}
\usepackage{amssymb,amsfonts}
\usepackage{multicol}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage[cm]{fullpage}
\newcommand{\tab}{\hspace*{5em}}
\newcommand{\conj}{\overline}
\newcommand{\dd}{\partial}
\newcommand{\ep}{\epsilon}
\newcommand{\openm}{\begin{pmatrix}}
\newcommand{\closem}{\end{pmatrix}}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\im}{im}
\DeclareMathOperator{\Span}{span}
\DeclareMathOperator{\Null}{null}
\newcommand{\nc}{\newcommand}
\newcommand{\rn}{\mathbb{R}}
\newcommand{\zn}{\mathbb{Z}}
\nc{\cn}{\mathbb{C}}
\nc{\ssn}[1]{\subsubsection*{#1}}
\nc{\inner}[2]{\langle #1,#2\rangle}
\nc{\h}[1]{\widehat{#1}}
\nc{\tl}[1]{\widetilde{#1}}
\nc{\norm}[1]{\left\|{#1}\right\|}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
\begin{document}
Name: Hall Liu

Date: \today 
\vspace{1.5cm}
\subsection*{1}
\ssn{a}
Suppose we have a population which is uniformly distributed on the unit circle. Consider some optimal codebook with two clusters. Since the population distribution has rotational symmetry, the population clustering risk is invariant under rotations. Thus, we can obtain a distinct codebook which is also optimal by rotating any optimal codebook. Since we can approach the optimal risk by approaching any of the rotated variants of the optimal codebook, we do not necessarily have to be close to the specified optimal codebook.
\ssn{b}
Let $C_k$ be the optimal clustering with $k$ clusters. If we arbitrarily add a center $d$ to this clustering to obtain a clustering $D$, we will have $R(D)\leq R(C_k)$. For any point $x$ in the space we're in, we have $\min_D\|x-d_i\|^2\leq\min_{C_k}\|x-c_i\|^2$, as $C_k$ is a subset of $D$. Thus, when we integrate to find the expected value, the integral with $D$ will be less than or equal by comparison, which means that we can always find a clustering with $k+1$ clusters that either matches or equals the optimal with $k$ clusters.

\end{document}

### 2014-04-16 stat246/hw2.tex
c2

If we work in terms of $\theta$, the likelihood for the data $X_i$ is 
\[\frac{\theta^{N\alpha}}{\Gamma(\alpha)^N}\prod_{i=1}^NX_i^{\alpha-1}e^{-X_i\theta}\]
Multiplying this by the conjugate prior produces
\[\left(\frac{\prod_{i=1}^NX_i^{\alpha-1}}{\Gamma(\alpha)^N}\cdot\frac{b^a}{\Gamma(a)}\right)\theta^{N\alpha+a-1}e^{-\theta\sum_{i=1}^NX_i-b\theta}\]
Up to some normalizing constant that absorbs all the stuff in the front in the parens, this is another gamma distribution with shape $a+N\alpha$ and rate $b+\sum X_i$, so the posterior mean of $\theta$ is $\frac{a+N\alpha}{b+\sum X_i}$, which means that the posterior mean of $\beta$ is the reciprocal of this.

>>>>

If we work in terms of $\theta$, the likelihood for the data $X_i$ is 
\[\frac{\theta^{N\alpha}}{\Gamma(\alpha)^N}\prod_{i=1}^NX_i^{\alpha-1}e^{-X_i\theta}\]
Multiplying this by the conjugate prior produces
\[\left(\frac{\prod_{i=1}^NX_i^{\alpha-1}}{\Gamma(\alpha)^N}\cdot\frac{b^a}{\Gamma(a)}\right)\theta^{N\alpha+a-1}e^{-\theta\sum_{i=1}^NX_i-b\theta}\]
Up to some normalizing constant that absorbs all the stuff in the front in the parens, this is another gamma distribution with shape $a+N\alpha$ and rate $b+\sum X_i$. Thus, $\beta=1/\theta$ has an inverse-gamma distribution with shape $a+N\alpha$ and scale $b+\sum X_i$, which means that the posterior mean of $\beta$ is $\frac{b+\sum X_i}{a+N\alpha-1}$
